Kronos \& Krios: TBD

### 2017-05-20 paper/ms.tex
Improve Tc rank vs delta X/H figure
working on summary of previous studies

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\tmass}{\project{\acronym{2MASS}}}
\newcommand{\documentname}{Article}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}
\newcommand{\dr}{\acronym{DR1}}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\etal}{\textit{et al}.}
\newcommand*\elem[1]{\ensuremath{\mathrm{#1}}}
\newcommand*\elemH[1]{\ensuremath{[\mathrm{#1}/\elem{H}]}}
\newcommand*{\feh}{\ensuremath{\elemH{Fe}}}
\newcommand{\sunanalog}{\text{Krios}}
\newcommand{\bizarreone}{\text{Kronos}}

@@

Combining these precise radial velocities with the \gaia\ astrometry, we can
compare differences between the inferred 6D phase-space properties for the
two stars.
We start by generating posterior samples over the Heliocentric distance, $r$,
tangential velocities, $(v_\alpha, v_\delta)$, and radial velocity, $\hat{v}_r$,
given the observed parallax, $\pi$, and proper motion components,
$(\mu_{\alpha^*}, \mu_\delta)$, and radial velocity, $v_r$.
By defining the vectors
% SMOH: I understand you meant hat v_r as the latent true radial velocity but
% personally think a) it can confuse some people that there are hat quantities
% with no-hat quantities for hat-y vector and thus b) it'd be better to use hat
% quantities for observed variables and no-hat for latent true ones.
% SMOH: v_r is missing in the LHS of eq(3)
\begin{eqnarray}
  \vec{y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          \pi &
          \mu_{\alpha^*} &
          \mu_\delta &
          v_r
        \end{array}
      \right)}\\
  \hat{\vec{y}} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          r^{-1} &
          r^{-1}\,v_\alpha &
          r^{-1}\,v_\delta &
          \hat{v}_r
        \end{array}
      \right)}
\end{eqnarray}
and by considering the covariance matrix $\mat{C}$ between the observed
quantities (provided by \tgas\ and extended to include the uncorrelated radial
velocity uncertainty), the likelihood can be written
\begin{eqnarray}
p(\pi, \mu_{\alpha^*}, \mu_\delta \given r, v_\alpha, v_\delta) &=&
  \left[\det\left(\frac{\mat{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\vec{y} - \hat{\vec{y}} \right)} \,
    \mat{C}^{-1} \,
    \left(\vec{y} - \hat{\vec{y}} \right) \right] \label{eq:likefn} \quad .
\end{eqnarray}
We adopt a uniform space density prior for the distance and an isotropic
Gaussian for any velocity component, $v$, with a dispersion $\sigma_v=25~\kms$
\begin{eqnarray}
p(r) &=&
  \begin{cases}
    \frac{3}{r_{\rm lim}^3} \, r^2 & \text{if } 0 < r < r_{\rm lim}\\
    0              & \text{otherwise}
  \end{cases}\\
p(v) &=& \frac{1}{\sqrt{2\pi}\,\sigma_v} \,
  \exp\left[-\frac{1}{2} \, \frac{v^2}{\sigma_v^2} \right] \quad .
\end{eqnarray}
For each of the two stars, we use \project{emcee}
(\citealt{Foreman-Mackey:2013}) to generate posterior samples in $(r, v_\alpha,
v_\delta, \hat{v}_r)$ by running 64 walkers for 4608 steps and discarding the
first 512 steps as the burn-in period.
For each sample, we convert the heliocentric phase-space coordinates into
Galactocentric coordinates assuming the Sun's position and velocity are $x_\odot
= (-8.3, 0, 0)~{\rm kpc}$ and $v_\odot = (10, 244, 7.17)~\kms$ (\citealt{bovy,
etc.}).
\figurename~\ref{fig:dxdv} shows differences in posterior samples converted to
Galactocentric phase-space coordinates for the two stars.
In all components, the pair is consistent with being co-moving and close in
separation; the sky-projected separation of the two stars is $\approx$$0.01~\pc$
but the median 3-space separation is $\approx$$0.6~\pc$.

@@

\figurename~\ref{fig:orbit} shows an orbit computed from the median posterior
sample over 6D phase-space coordinates for \sunanalog\ (black line), integrated
in a Milky Way-like gravitational potential (\cite{Gala:2017}). \todo{APW:
explain}
The Sun's orbit in this same Milky Way model is under-plotted for comparison
(grey line); the orbits of the stars in this pair have large scale-heights
relative to the Sun's orbit despite being young.

@@

Given their proximity in space and kinematics,
it is generally accepted to assume that the two stars were born coeval
perhaps in a wide binary from the same birth cloud.
Thus, we expect the two stars to have identical metallicities and abundance patterns,
\todo{except for thoese elements like Li...}
Surprisingly, one of the stars, \bizarreone\ is significantly more metal
rich than the other by 0.2~dex ($\approx 60\%$; Figure~\ref{fig:abundances}).
What is more puzzling is that the abundances of this star
shows selective depletion in C, N, O, Na, and Mn.

@@

Surface lithium abundance in a sun-like star decreases with its age
due to mixing induced by convection or rotation which brings lithium
into the interior ($T>2.5 \times 10^{6}$~K)
where it will be destroyed by proton capture burning.
Thus, surface lithium abundance is an indicator of stellar ages.
The $\elem{Li}$ abundance of $2.25$ for \sunanalog\ implies an age of $\lesssim 1$~Gyr
according to the theoretical models tuned to explain the solar lithium abundance
and rotational profile (\citealt{2005Sci...309.2189C}).
The lithium abundance of \bizarreone\ is still higher by $0.5$~dex
showing the largest difference among all measured elements.
This translates to $\sim 500$~Myr difference in age.
Given the overall higher metal abundances and the peculiar abundance patterns in \bizarreone,
it is unclear, however, whether this higher $\elem{Li}$ abundance
means a younger age or something else.
For example, the presence of $\elem{Li}$-rich red giant stars has been attributed
to the engulfment of substellar companions such as gas giant planets or brown dwarfs
which may replenish $\elem{Li}$ (\citealt{Casey:2016aa}).

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.9\linewidth]{abundances.pdf}
  \caption{Abundances of the co-moving pair, \sunanalog\ and \bizarreone\.}
  \label{fig:abundances}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.9\linewidth]{relabun_tc_XH.pdf}
  \caption{Abundance of \bizarreone\ ($[\mathrm{Fe}/\mathrm{H}] = 0.20$)
    relative to \sunanalog\ ($[\mathrm{Fe}/\mathrm{H}] = 0.01$)
    as a function of condensation temperature of elements (\citealt{2003ApJ...591.1220L}).
  }
  \label{fig:relabun_tc}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[angle=90,height=0.9\textheight]{TcRank_deltaXH.pdf}
  \caption{Abundance of \bizarreone\ ($[\mathrm{Fe}/\mathrm{H}] = 0.20$)
    relative to \sunanalog\ ($[\mathrm{Fe}/\mathrm{H}] = 0.01$)
    as a function of condensation temperature of elements (\citealt{2003ApJ...591.1220L}).
  }
  \label{fig:relabun_tcrank}
\end{figure}

@@

None of the above scenarios is satisfactory as it offers no
insight on the peculiarity in abundances of HD 30 (see \sectionname~{\ref{sec:data}).

The unlikeliness of the above possibilities naturally leads us to
consider the alternative.
What can selectively enrich XX elements in one of the two stars that are born together?

% inhomogeneity in star formation
It may be that the two stars are born at the same time from the same cloud, but
there is inhomogeneity within the cloud.
However, there is already ample amount of evidence against such scenario.
First, none of the other seven similar wide binaries examined in \citealt{2016ApJS..225...32B}
show the same level differences in abundances although there is generally
a larger spread in $\elem{C}$, $\elem{N}$ and $\elem{O}$, and some pairs show
a large as $\approx 0.15$~dex difference in particular elements.
The median and maximum \feh\ difference between component stars in the other seven pairs
is $0.02$~dex and $0.09$~dex, respectively.
This is consistent with the findings of \citealt{Desidera:2004aa} who examined 23 wide bianries
of late F to K dwarfs, and found most pairs show difference in \feh\ less than $0.02$~dex
and none larger than $0.07$~dex.

@@

% rocky planets (or the like) engulfment
Yet another possibility is that the more metal-rich of the two stars
was recently? bombarded by something that selectively removed the depleted elements but not the others.

@@

Change in stellar abundance due to ingestion of planets has been considered
for similar bright star pairs HD XX/XX (Mack) and HD XX/XX(Mack).

% Concluding remakrs

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{spec1.pdf}
  \caption{Selective segments of spectra of \sunanalog\ and \bizarreone.
    For each row, the continuum-normalized data and model are plotted in the upper panel,
    and the ratio (\bizarreone/\sunanalog) of data (gray) and model (black)
    are plotted in the lower panel.
  }
  \label{fig:spec1}
\end{figure}

>>>>

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\tmass}{\project{\acronym{2MASS}}}
\newcommand{\documentname}{Article}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}
\newcommand{\dr}{\acronym{DR1}}
\newcommand{\tgas}{\acronym{TGAS}}
\newcommand{\etal}{\textit{et al}.}
\newcommand*\elem[1]{\ensuremath{\mathrm{#1}}}
\newcommand*\elemH[1]{\ensuremath{[\mathrm{#1}/\elem{H}]}}
\newcommand*{\feh}{\ensuremath{\elemH{Fe}}}
\newcommand{\sunanalog}{\text{Krios}}
\newcommand{\bizarreone}{\text{Kronos}}
\newcommand{\Tcondens}{\ensuremath{T_C}}
\newcommand{\mearth}{\ensuremath{M_\oplus}}
\newcommand{\mjupiter}{\ensuremath{M_\mathrm{Jup}}}

@@

Combining these precise radial velocities with the \gaia\ astrometry, we can
compare differences between the inferred 6D phase-space properties for the
two stars.
We start by generating posterior samples over the Heliocentric distance, $r$,
tangential velocities, $(v_\alpha, v_\delta)$, and radial velocity, $\hat{v}_r$,
given the observed parallax, $\pi$, and proper motion components,
$(\mu_{\alpha^*}, \mu_\delta)$, and radial velocity, $v_r$.
By defining the vectors
% SMOH: I understand you meant hat v_r as the latent true radial velocity but
% personally think a) it can confuse some people that there are hat quantities
% with no-hat quantities for hat-y vector and thus b) it'd be better to use hat
% quantities for observed variables and no-hat for latent true ones.
% SMOH: v_r is missing in the LHS of eq(3)
\begin{eqnarray}
  \vec{y} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          \pi &
          \mu_{\alpha^*} &
          \mu_\delta &
          v_r
        \end{array}
      \right)}\\
  \hat{\vec{y}} &=&
      \transp{\left(
        \begin{array}{c@{\hspace{1em}} c@{\hspace{1em}} c@{\hspace{1em}} c}
          r^{-1} &
          r^{-1}\,v_\alpha &
          r^{-1}\,v_\delta &
          \hat{v}_r
        \end{array}
      \right)}
\end{eqnarray}
and by considering the covariance matrix $\mat{C}$ between the observed
quantities (provided by \tgas\ and extended to include the uncorrelated radial
velocity uncertainty), the likelihood can be written
\begin{eqnarray}
p(\pi, \mu_{\alpha^*}, \mu_\delta \given r, v_\alpha, v_\delta) &=&
  \left[\det\left(\frac{\mat{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\vec{y} - \hat{\vec{y}} \right)} \,
    \mat{C}^{-1} \,
    \left(\vec{y} - \hat{\vec{y}} \right) \right] \label{eq:likefn} \quad .
\end{eqnarray}
We adopt a uniform space density prior for the distance and an isotropic
Gaussian for any velocity component, $v$, with a dispersion $\sigma_v=25~\kms$
\begin{eqnarray}
p(r) &=&
  \begin{cases}
    \frac{3}{r_{\rm lim}^3} \, r^2 & \text{if } 0 < r < r_{\rm lim}\\
    0              & \text{otherwise}
  \end{cases}\\
p(v) &=& \frac{1}{\sqrt{2\pi}\,\sigma_v} \,
  \exp\left[-\frac{1}{2} \, \frac{v^2}{\sigma_v^2} \right] \quad .
\end{eqnarray}
For each of the two stars, we use \project{emcee}
(\citealt{Foreman-Mackey:2013}) to generate posterior samples in $(r, v_\alpha,
v_\delta, \hat{v}_r)$ by running 64 walkers for 4608 steps and discarding the
first 512 steps as the burn-in period.
For each sample, we convert the heliocentric phase-space coordinates into
Galactocentric coordinates assuming the Sun's position and velocity are $x_\odot
= (-8.3, 0, 0)~{\rm kpc}$ and $v_\odot = (10, 244, 7.17)~\kms$ (\citealt{bovy,
etc.}).
\figname~\ref{fig:dxdv} shows differences in posterior samples converted to
Galactocentric phase-space coordinates for the two stars.
In all components, the pair is consistent with being co-moving and close in
separation; the sky-projected separation of the two stars is $\approx$$0.01~\pc$
but the median 3-space separation is $\approx$$0.6~\pc$.

@@

\figname~\ref{fig:orbit} shows an orbit computed from the median posterior
sample over 6D phase-space coordinates for \sunanalog\ (black line), integrated
in a Milky Way-like gravitational potential (\cite{Gala:2017}). \todo{APW:
explain}
The Sun's orbit in this same Milky Way model is under-plotted for comparison
(grey line); the orbits of the stars in this pair have large scale-heights
relative to the Sun's orbit despite being young.

@@

Given their proximity in space and kinematics,
it is generally accepted to assume that the two stars were born coeval
perhaps in a wide binary from the same birth cloud.
Thus, we expect the two stars to have identical metallicities and abundance patterns,
\todo{except for thoese elements like Li...}
Surprisingly, one of the stars, \bizarreone\ is significantly more metal
rich than the other by 0.2~dex ($\approx 60\%$; \figname~\ref{fig:abundances}).
What is more puzzling is that the abundances of this star
shows selective depletion in C, N, O, Na, and Mn.

@@

Surface lithium abundance in a sun-like star decreases with its age
due to mixing induced by convection or rotation, which brings the lithium
into the interior ($T>2.5 \times 10^{6}$~K)
where it will be destroyed by proton capture burning.
Thus, surface lithium abundance is an indicator of stellar ages.
The $\elem{Li}$ abundance of $2.25$ for \sunanalog\ implies an age of $\lesssim 1$~Gyr
according to the theoretical models tuned to explain the solar lithium abundance
and rotational profile (\citealt{2005Sci...309.2189C}).
The lithium abundance of \bizarreone\ is still higher by $0.5$~dex
showing the largest difference among all measured elements.
This translates to $\sim 500$~Myr difference in age.
Given the overall higher metal abundances and the peculiar abundance patterns in \bizarreone,
it is unclear, however, whether this higher $\elem{Li}$ abundance
means a younger age or something else.
For example, the presence of $\elem{Li}$-rich red giant stars has been attributed
to the engulfment of substellar companions such as gas giant planets or brown dwarfs
which may replenish $\elem{Li}$ (\citealt{Casey:2016aa}).

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.9\linewidth]{abundances.pdf}
  \caption{Abundances of the co-moving pair, \sunanalog\ and \bizarreone.}
  \label{fig:abundances}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{TcRank_deltaXH_concise.pdf}
  \caption{Abundance of \bizarreone\ ($[\mathrm{Fe}/\mathrm{H}] = 0.20$)
    relative to \sunanalog\ ($[\mathrm{Fe}/\mathrm{H}] = 0.01$)
    as a function of condensation temperature of elements (\citealt{2003ApJ...591.1220L}).
  }
  \label{fig:relabun_tcrank}
\end{figure}

@@

None of the above scenarios is satisfactory as they offer no
insight on the peculiarity of the abundances of \bizarreone
(see \sectionname~{\ref{sec:data}}).
In case that the pair formed from a exchange scattering event,
we would expect the pair to be composed of typical field stars.
The unlikeliness of the above possibilities naturally leads us to
consider the alternative.
What can selectively enrich refractory (high $T_C$) elements
in one of the two stars that are born together?

% inhomogeneity in star formation
It may be that the two stars are born at the same time from the same cloud, but
there is inhomogeneity within the cloud.
However, there is already ample amount of evidence against such scenario.
First, none of the other seven similar wide binaries examined in \citealt{2016ApJS..225...32B}
show the same level differences in abundances although there is generally
a larger spread in $\elem{C}$, $\elem{N}$ and $\elem{O}$, and some pairs show
a large as $\approx 0.15$~dex difference in particular elements.
The median and maximum \feh\ difference between component stars in the other seven pairs
is $0.02$~dex and $0.09$~dex, respectively.
This is consistent with the findings of \citealt{Desidera:2004aa}, who examined 23 wide binaries
of late F to K dwarfs, and found most pairs show difference in \feh\ less than $0.02$~dex
and none larger than $0.07$~dex.

@@

% rocky planets (or the like) engulfment
Yet another possibility that may change a star's surface abundance after formation
is accretion of its planets or asteroids.
Instabilities may develop in a multi-planet system due to chaotic
nature of many body system, which may lead to planet engulfment by the host star.

One informative proxy that may reveal the composition of the hypothetical material
that was accreted to the stars is the condensation temperature.

@@

\citealt{Mack:2014aa} studied HD~20782/20781, two common proper motion G dwarf stars (G2/G9.5)
with a projected separation of $\sim9000$~AU (correspoding to 4.2~arcmin sky separation).
Both stars in this system host close-in giant planets.
HD~20782 hosts a Jupiter-mass planet on a very eccentric ($e\approx 0.97$) orbit with a
pericenter distance of 1.4~AU while HD~20781 hosts two Neptune-mass planets within 0.3~AU
with moderately high eccentricity ($e\sim0.1-0.3$).
They found that the abundances of 15 elements between the two stars are consistent with each other.
However, they argued that there is a moderately significant ($\sim 2\sigma$)
positive slope of $\approx 10^{-5}$~dex\,K$^{-1}$ in abundance of each star
{\it individually} with increasing \Tcondens\ for $\Tcondens>900$~K elements
(namely, Na, Mn, Cr, Si, Fe, Mg, Co, Ni, V, Ca, Ti, Al, Sc leaving out C and O
of their measurements).
They interpreted that this slope may be evidence for accretion of $10-20$~\mearth\ of 
H-depleted rocky material that were pushed in during giant planet migration.

\citealt{Teske:2016aa} studied WASP-94A/WASP-94B, a pair of F8 and F9 stars
with super-solar metallicity ($\feh\approx 0.3$ for both stars). Both stars
host a hot Jupiter.
The planet hosted by WASP-94A is transiting with a misaligned, probably
retrograde circular ($e<0.13$) orbit, while that hosted by WASP-94B is a little
more massive by $\sim 0.15$~\mjupiter\ and closer in, aligned with the host
star.
With a claimed precision of $0.006$~dex (median among all elements), they
detected a depletion of volatile and moderately volatile ($\Tcondens < 1200$~K)
elements at $0.02$~dex level and an enhancement in refractory ($\Tcondens >
1200$~K) elements at $0.01$~dex level in WASP-94A relative to WASP-94B.
Excluding K, which is based on one strong line and largely affected by non-LTE
effect, their analysis shows a statistically significant non-zero slope between
$\Delta\elemH{X}$ and $\Tcondens$\footnote{Note that the condensation
  temperature $\Tcondens$ used is for solar (system? atmosphere?) composition
  gas, which can differ from that of higher metallicity gas.}.
They considered multiple possibilities related to the formation and evolution
of planetary systems around each star as well as causes unrelated to planets
such as dust cleansing during the fully convective phase or different rotation
and granulation between the stars, but did not favor any particular one.

The pair of G5 stars HD 80606/HD 80607 has been independently studied by
\citealt{Saffe:2015aa} and \citealt{Mack:2016aa} using the same KECK HIRES
spectra. HD 80606 hosts a giant planet with mass of $4~\mjupiter$ at $0.5$~AU
on a very eccentric ($e\approx0.95$) orbit while HD 80607 has no detected
planets.
Both stars are super solar with $\feh=0.35$.

% Concluding remarks

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{spec1.pdf}
  \caption{Selective segments of the spectra of \sunanalog\ and \bizarreone.
    For each row, the continuum-normalized data and model are plotted in the upper panel,
    and the ratio (\bizarreone/\sunanalog) of data (gray) and model (black)
    are plotted in the lower panel.
  }
  \label{fig:spec1}
\end{figure}
